Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-06-13 cmsb22.tex
Add descriptions for using mpbn

However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
And hence, it is not possible to extract any (complex) attractor at all.
This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model or even of the starting map.

@@

\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.}

@@

We used a set of real-world Boolean models lying in various scales collected from numerous bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
Table~\ref{tab:result_real} shows the experimental results on those models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.

>>>>

However, using a state-of-the-art tool like PyBoolNet~\cite{klarner2015computing} on that model actually fails at the phase of prime-implicant generation.
\revise{Note that mpbn does not give any answer because that model is not locally-monotonic.}
And hence, it is not possible to extract any (complex) attractor at all.
This is also true for the Alzheimer model also mentioned in that same article and originally from~\cite{ogishima2016alzpathway} (\verb|F4| file in the original supplementary material, and ``Alzheimer'' in Table~\ref{tab:result_real}), but actually not for the MAPK model for which the first trap spaces can be obtained in reasonable time.
The current practice usually revolves then around fixing some inputs to plausible values and reducing the model accordingly.
While this approach makes sense, it relies on potentially arbitrary decisions, and \emph{hides away} critical modelling choices that were actually not part of the original Boolean model or even of the starting map.

@@

\revise{To assess the efficiency of the proposed method, implemented as a Python package named Trappist, we compare it with the state-of-the-art method implemented in the tool PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet} on both its own repository of models and large models from the literature.
In addition, we also include the tool mpbn~\cite{Paulev2020}\footnote{We use the latest version of mpbn available at submission time through CoLoMoTo.} to the benchmarks although it only handle locally-monotonic models, whereas both Trappist and PyBoolNet can handle general models.}

@@

We used a set of real-world Boolean models lying in various scales collected from numerous bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
\revise{We then applied PyBoolNet, mpbn, and Trappist} to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
Table~\ref{tab:result_real} shows the experimental results on those models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.
